Alleviating the Topology Mismatch Problem in Distributed Overlay Networks

### 2014-01-28 conclusion.tex
2013.01.23 (Vassilis) Text refinements in introduction and conclusion sections.

Research and development for computer systems
built using the \p\ paradigm have flourished  in the last decade.
By and large, such systems exploit the loose-coupling and self-organization 
of participating nodes to shape substrates upon which diverse applications 
can run; the latter offer services including but not limited to 
distributed naming, 
Internet telephony and video conferencing, 
storage and indexing, 
content and file sharing, 
web crawling and caching, 
event notification and subscribe/publish.
These services are provided atop the 
best-effort infrastructure of today's Internet 
and they have to effectively deal with an array of key issues
emanating from the fact that \p\ nodes 
operate at the edge of the Internet; such critical issues include
node availability, robustness, scalability, load-sharing,
quality-of-service and user anonymity.
The majority of such issues emerges as 
application-level \p--networks are formed without paying 
adequate attention to the design of the physical network beneath.
The latter constitutes what is known as the
topology mismatch problem between the \p-overlays and the physical networks.
In addition, the development of \p--systems and applications
typically involved juggling a set of tradeoffs (e.g.,
choosing higher accuracy at the cost of increased system overhead.)

% Computer systems design is typically about juggling a set of tradeoffs (e.g.,
% choosing higher accuracy at the cost of increased system overhead.)  
In surveying more than
a decade's worth of research efforts aimed at solving the topology mismatch
problem in both unstructured and structured \p\ networks, we find this
tussle-of-tradeoffs property to hold.  
With regards to the three criteria we used --efficiency,
overhead, and scalability-- we find that none of the proposed solutions 
is superior to the others on all three fronts.  This is not surprising.
Nonetheless, by
1)~presenting an analysis of each of the solutions, 
including what distinguishes it from the
others and its advantages and disadvantages and 
2)~offering a pictorial
comparison of how each solution fares compared with the others with regards
to efficiency, overhead, and scalability, we hope to have provided \p\
researchers and developers with enough perspective so that depending 
on the needs of their particular application, they are able to select
amongst these surveyed solutions or propose a new solution of their own.

>>>>

Research and development around application systems that follow the 
\p\ architectural paradigm, have flourished in the last ten years or so.
By and large, such systems exploit the loose-coupling and self-organization 
of participating nodes, to shape substrates upon which diverse applications 
can run; the latter offer services including, but not limited to, 
distributed naming, 
Internet telephony and video conferencing, 
storage and indexing, 
content and file sharing, 
web crawling and caching, 
event notification and subscribing/publishing.
These services are provided atop the 
best-effort infrastructure of today's Internet 
and they have to effectively deal with an array of key quality attributes that
they may need or have to offer, such as node availability, robustness,
scalability, load-sharing, quality-of-service and user anonymity.
The difficulty to handle such features effectively, emanates from the fact that
peer nodes operate at the edge of the Internet and no adequate attention to the
structure of the physical network is paid during application-level network
formation. The deviation of the structure of the overlay network from the
optimal, according to the underlying physical structure, constitutes what is
termed topology mismatch and the problem is known as the topology mismatch
problem.

Development of \p--systems and applications
typically involve juggling a set of tradeoffs (e.g.,
choosing higher accuracy at the cost of increased system overhead.).
% Computer systems design is typically about juggling a set of tradeoffs (e.g.,
% choosing higher accuracy at the cost of increased system overhead.)  
In surveying more than
a decade's worth of research efforts aimed at solving the topology mismatch
problem in both unstructured and structured \p\ networks, we find this
tussle-of-tradeoffs property to hold.  
With regards to the three criteria we used --efficiency,
overhead, and scalability-- we find that none of the proposed solutions 
is superior to the others on all three fronts.  This is not surprising.
Nonetheless, by
\begin{inparaenum}[\itshape i\upshape)]
  \item presenting an analysis of each of the solutions,
including what distinguishes it from the
others and its advantages and disadvantages and
  \item offering a pictorial
comparison of how each solution fares compared with the others with regards
to efficiency, overhead, and scalability,
\end{inparaenum}
we hope to have provided \p\
developers and researchers with enough perspective so that depending
on the particular problem in hand, they are able to make a
correct choice amongst these surveyed here or, by firmly standing on
the shoulders of giants, propose a new solution of their own.

### 2014-01-28 intro.tex
2013.01.23 (Vassilis) Text refinements in introduction and conclusion sections.

\emph{Peer-to-peer (\p)} networks are self-organizing distributed systems where
participating nodes, called \emph{peers}, act as both resource providers and
resource consumers in contrast to the traditional \emph{client-server} model
where nodes undertake specific roles.
For over a decade, \p\ networks have been widely deployed and have
enjoyed immense popularity from Internet communities, primarily because
of the great number of features they offer to distributed applications 
built atop them. 
Such diverse features include:  high availability and robustness,
load-balancing, quality of service, scalability, decentralized administration,
and anonymity.

The peer-to-peer paradigm gave impetus to two ``killer'' applications:
file-sharing and Internet telephony.
The {\sl Napster} file-sharing was widely acknowledged as the 
``fastest growing Internet application ever" in $2001$ when 
$26$ million users shared $80$ million songs.
% File-sharing, originally introduced by
% {\sl Napster} (Section~\ref{section:background}), was widely
% acknowledged by $2001$ as the ``fastest growing Internet application ever" 
% with $26$ million users sharing $80$ million songs.
% % at the height of its popularity. 
Other file-sharing applications followed suit 
including {\sl Gnutella} and {\sl Limewire} enjoying $3$ 
million concurrently-connected peers as well as 
{\sl BitTorrent} connecting over $150$ million users by the end of $2006$.
\p\ telephone saw explosive growth with the advent of {\sl Skype}. 
Since its introduction in $2003$,
{\sl Skype} has become extremely popular with more than $650$ million users 
in $2011$~\cite{skypetotalusers}; on March $5$th, $2012$,  
the astonishing figure of $35$ million concurrently-online users  
was reported~\cite{skypesymusers}.
%%
In addition, a startling number of diverse and successful
applications have been built on top of \p\ systems including:
distributed search engines~\cite{yaci}, 
distributed data-storage systems~\cite{kbc_oceanstore_2000,bdet_fsdfs_2000,dkkms_cfs_2001,dr_pastutility_2001,abc_farsite_2002,mmfc_ivy_2002,arla,agebh_dks_2003},
Web caches, archives and publishing systems~\cite{ird_squirrel_2002,bags_youserv_2002,wrc_publius_2000,wm_tangler_2001},
messaging applications~\cite{threedegrees}, 
event-notification infrastructures~\cite{rkcd_scribe_2001,cdkr_scribe_2002,agebh_dks_2003}, 
naming services~\cite{cmm_chorddns_2002}, 
censor-resistant stores~\cite{cswh_freenet_2001} and
lately, even cloud-based platforms~\cite{mgpj_cloudsnap_2011}.

\p\ systems are implemented using \emph{overlay networks}. 
An overlay network is a virtual system of nodes and logical links
created atop an existing network; overlays provide 
added value to underlying network by for instance offering QoS
content-distribution over the best effort Internet. 
Overlay nodes communicate through \emph{virtual connections} each 
of which may correspond to a path of possibly many physical links 
in the underlying network.
Figure~\ref{figure:overlay} illustrates a simple four-node overlay constructed
over a wide-area network.
%%
\begin{figure}[ht]
\centering
  \includegraphics[scale=0.6]{img/pdf/under-over-lay.pdf}
\caption{An example overlay network.}
\label{figure:overlay}
\end{figure}

The single key issue that determines the efficiency of an overlay network
is how well the overlay maps to the underlying network topology on which it
``rests''. 
Consider two nodes\footnote{In \p\ networks,
the participating nodes are typically user-PCs operating at the edge of the
Internet.} that share a overlay network link which ultimately 
maps to multiple hops in the underlying IP topology. 
If the application running on the nodes generates traffic flows 
the heavily use the link in question, it would be beneficial to 
construct the overlay topology in a way that the number of 
underlying IP links between these two nodes is minimized. 
% If the traffic patterns of the application
% are such that traffic flows heavily over the overlay link 
% between the two nodes,
% then it is beneficial to construct the overlay 
% topology such that the number of
% underlying IP links between these two nodes is minimized.
%%
Should the overlay network be constructed  so that 
it does not match the underlying topology well, 
the inherent \emph{topology mismatch} gives two
significant problems: first, the performance of the application per se, can be
adversely affected since traffic must flow over a larger,
redundant, number of physical hops resulting in poor user experience 
entailing noticeable latencies or jittering. 
%%
Second, other applications running
on the underlying network infrastructure are also adversely affected.
Studies have shown that highly popular \p\ applications contribute 
the largest portion of overall 
Internet traffic~\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with some reporting that more than $60$\% of this traffic 
to be \p-related~\cite{cachelogic,ipoque2007,ipoque2009};
it was also projected that this traffic would reach
$8\;$\emph{Petabytes}-per-month by the end of $2012$~\cite{multinteligence}. 
Evidently, this constitutes a major burden for 
Internet Service Providers (ISPs) who must route 
all of this traffic to destinations at the edge of the Internet. 
If the \p\ overlay topology is poorly designed, 
the demand on the Internet's backbone infrastructure may 
substantially increase as traffic might have to flow
``back and forth'' several times between two neighboring ISPs 
while trying to travel from the source to its destination node
in the overlay.
Hence, it is very critical that \p\ networks be laid out 
in ways that their topology matches as closely as possible the 
underlying IP topology.

For over a decade, researchers have extensively investigated  
various aspects of this topology mismatch problem.  
The main objective of this paper is to offer a comprehensive 
survey of the work done in this are and provide a
taxonomy of the proposed solutions. We point out synergies as well as
similarities and differences in the published approaches. 
Ultimately, our goal is to help readers sift through 
the voluminous literature, to help them
understand the advantages and disadvantages of each work, and 
to provide them with enough perspective so that 
when the need arises, they are able to
select amongst the different approaches for their particular application.
%%
The rest of the paper is organized as follows: 
Section~\ref{section:background} provides background on
overlay architectures including centralized, decentralized
unstructured and decentralized-structured \p\ systems. 
We also formally define the problem of topology mismatch 
% between an overlay and its underlying
% physical network 
and offer the rationale behind our 
evaluation of the techniques discussed in this paper. 
Sections~\ref{section:unstructured} and~\ref{section:structured}
respectively outline the surveyed research efforts 
for decentralized-unstructured and decentralized-structured \p\ systems.
Conclusions can be found in Section~\ref{section:conclusion}.

>>>>

\emph{Peer-to-peer} (\p) networks are self-organizing, distributed systems where
participating nodes, called \emph{peers}, act as both resource providers and
resource consumers, in contrast to the traditional \emph{client-server} model
where nodes undertake specific roles.
For over a decade, \p\ networks have been widely deployed and have
enjoyed immense popularity from Internet communities, primarily because
of the great number of features they offer to distributed applications 
built atop them. 
Such diverse features include:  high availability and robustness,
load-balancing, quality of service, scalability, decentralized administration,
and anonymity.

The peer-to-peer paradigm gave impetus to two ``killer'' applications:
file-sharing and Internet telephony.
The {\sl Napster} file-sharing was widely acknowledged as the 
``fastest growing Internet application ever'' in $2001$ when
$26$ million users shared $80$ million songs.
% File-sharing, originally introduced by
% {\sl Napster} (Section~\ref{section:background}), was widely
% acknowledged by $2001$ as the ``fastest growing Internet application ever" 
% with $26$ million users sharing $80$ million songs.
% % at the height of its popularity. 
Other file-sharing applications followed suit, 
including {\sl Gnutella} and {\sl Limewire} enjoying $3$ 
million concurrently-connected peers as well as 
{\sl BitTorrent} connecting over $150$ million users by the end of $2006$.
\p\ telephony saw explosive growth with the advent of {\sl Skype}.
Since its introduction in $2003$,
{\sl Skype} has become extremely popular with more than $650$ million users 
in $2011$~\cite{skypetotalusers}; on March $5$th, $2012$,  
the astonishing figure of $35$ million concurrently-online users  
was reported~\cite{skypesymusers}.
%%
In addition, a startling number of diverse and successful
applications have been built based on a \p\ architecture. Some of them include:
distributed search engines~\cite{yaci}, 
distributed data-storage systems~\cite{kbc_oceanstore_2000,bdet_fsdfs_2000,dkkms_cfs_2001,dr_pastutility_2001,abc_farsite_2002,mmfc_ivy_2002,arla,agebh_dks_2003},
Web caches, archives and publishing systems~\cite{ird_squirrel_2002,bags_youserv_2002,wrc_publius_2000,wm_tangler_2001},
messaging applications~\cite{threedegrees}, 
event-notification infrastructures~\cite{rkcd_scribe_2001,cdkr_scribe_2002,agebh_dks_2003}, 
naming services~\cite{cmm_chorddns_2002}, 
censor-resistant stores~\cite{cswh_freenet_2001} and
lately, even cloud-based platforms~\cite{mgpj_cloudsnap_2011}.

\p\ systems are implemented using \emph{overlay networks}. 
An overlay network, is a virtual system of nodes and logical links
created above an existing network; overlays provide
added value to the underlying layer by, for instance, offering QoS
content-distribution over the best effort Internet. 
Overlay nodes communicate through \emph{virtual connections} each 
of which may correspond to a path of possibly many physical links 
in the underlying network.
Figure~\ref{figure:overlay} illustrates a simple four-node overlay constructed
over a wide-area network.
%%
\begin{figure}[ht]
\centering
  \includegraphics[scale=0.6]{img/pdf/under-over-lay.pdf}
\caption{An example overlay network.}
\label{figure:overlay}
\end{figure}

The single key issue that determines the efficiency of an overlay network,
is how well the overlay maps to the underlying network topology on which it
``rests''. 
Consider two nodes\footnote{In \p\ networks,
the participating nodes are typically user-PCs operating at the edge of the
Internet.} that are connected with each other via a path of overlay links.
If the application running on the nodes, generates heavy traffic along
the overlay path, it would be beneficial to
construct the overlay topology in a way that the number of 
underlying IP links between these two nodes is minimized.
% If the traffic patterns of the application
% are such that traffic flows heavily over the overlay link 
% between the two nodes,
% then it is beneficial to construct the overlay 
% topology such that the number of
% underlying IP links between these two nodes is minimized.
%%
Should the overlay network be constructed so that
it does not match the underlying topology well, 
the inherent \emph{topology mismatch} gives two
significant problems: first, the performance of the application per se, can be
adversely affected since traffic must flow over a larger,
redundant, number of physical hops resulting in poor user experience 
entailing noticeable latencies or jittering. 
%%
Second, other applications running
on the underlying network infrastructure are also adversely affected.
Studies have shown that highly popular \p\ applications contribute 
the largest portion of the overall 
Internet traffic~\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with some reporting that more than $60$\% of this traffic 
to be \p-related~\cite{cachelogic,ipoque2007,ipoque2009};
it was also projected that this traffic would reach
$8\;$\emph{Petabytes}-per-month by the end of $2012$~\cite{multinteligence}. 
Evidently, this constitutes a major burden for 
Internet Service Providers (ISPs) who must route 
all of this traffic to destinations at the edge of the Internet. 
If the \p\ overlay topology is poorly designed, 
the demand on the Internet's backbone infrastructure may 
substantially increase as traffic might have to flow
``back and forth'' several times between two neighboring ISPs
while trying to travel from the source to its destination node
in the overlay.
Hence, it is very critical that \p\ networks be laid out 
in ways that their topology matches, as closely as possible, the 
underlying IP topology.

For over a decade, researchers have extensively investigated  
various aspects of the topology mismatch problem.
The main objective of this paper is to offer a comprehensive 
survey of the work done in this and provide a
taxonomy of the proposed solutions. We point out synergies, as well as
similarities and differences in the published approaches. 
Ultimately, our goal is to help readers sift through 
the voluminous literature, to help them
understand the advantages and disadvantages of each work, and 
to provide them with enough perspective so that 
when the need arises, they are able to
select, amongst the different approaches, the one that is most suitable for
their particular application.
%%
The rest of the paper is organized as follows: 
Section~\ref{section:background} provides background on
overlay architectures including centralized, decentralized
unstructured and decentralized-structured \p\ systems. 
We also formally define the problem of topology mismatch 
% between an overlay and its underlying
% physical network 
and offer the rationale behind our 
evaluation of the techniques discussed in this paper. 
Sections~\ref{section:unstructured} and~\ref{section:structured}
respectively outline the surveyed research efforts 
for decentralized-unstructured and decentralized-structured \p\ systems.
Conclusions can be found in Section~\ref{section:conclusion}.

### 2014-01-28 mard-survey.tex
2013.01.23 (Vassilis) Text refinements in introduction and conclusion sections.

\begin{abstract}
Peer-to-peer (\p) systems have enjoyed immense attention and have been
widely deployed on the Internet for well over a decade.  
They are implemented via an overlay network abstraction on top 
of the best-effort IP infrastructure of the Internet. 
\p\ systems offer a plethora of desirable features to 
distributed applications that are built on top of them including
anonymity, high availability and robustness, 
load balancing, quality of service, scalability, and others.  
Unfortunately, weaknesses of early deployments of \p\ systems prevented 
applications from leveraging these features.  
One major weakness, identified early on, is the topology
mismatch problem between the overlay network and the underlying IP topology.
This mismatch can impose an extraordinary amount of unnecessary stress on
network resources and can adversely affect the scalability of applications
built atop \p\ systems.  
In this paper, we survey over a decade's worth of research efforts
aimed at alleviating the topology mismatch problem, in both unstructured
and structured \p\ systems.   
We provide a fine-grained categorization of the
proposed solutions and discuss the advantages, disadvantages, and novelties
introduced by each solution.  
Finally, we provide an analysis as well as pictorial comparisons 
of the proposed solutions since we aim to offer a comprehensive 
reference for system architects, developers, and researchers in the field.
% in an attempt to provide a comprehensive quick one-stop-shop
% reference for anyone interested in the topic, from researchers and students,
% to system architects and developers.

@@

\begin{bottomstuff}
Authors' addresses: Deprt. of Informatics and Telecommunications,
University of Athens, Athens, Greece. Deprt. of Software
Engineering, Izmir University of Economics, Izmir, Turkey.\\
emails: \{b.moystakas,mema,ad\}@di.uoa.gr, huseyin.akcan@ieu.edu.tr
\end{bottomstuff}

>>>>

\begin{abstract}
\emph{Peer-to-peer} (\p) systems have enjoyed immense attention and have been
widely deployed on the Internet for, well, over a decade.  
They are implemented via an overlay network abstraction, on top 
of the best-effort IP infrastructure of the Internet. 
\p\ systems offer a plethora of desirable features to 
distributed applications that are built on top of them including
anonymity, high availability and robustness, 
load balancing, quality of service, scalability, to name just a few.
Unfortunately, weaknesses of early deployments of \p\ systems, prevented 
applications from leveraging these features.
One major weakness, identified early on, is the \emph{topology mismatch problem}
between the overlay network and the underlying IP topology.
This mismatch can impose an extraordinary amount of unnecessary stress on
network resources and can adversely affect the scalability and efficiency of
applications built atop.
In this paper, we survey over a decade's worth of research efforts
aimed at alleviating the topology mismatch problem, in both unstructured
and structured \p\ systems.   
We provide a fine-grained categorization of the
proposed solutions and discuss the advantages, disadvantages, and novelties
introduced by each of them.
Finally, we provide an analysis as well as pictorial comparisons 
of the reviewed approaches, since we aim to offer a comprehensive
reference for developers, system architects and researchers in the field.
% in an attempt to provide a comprehensive quick one-stop-shop
% reference for anyone interested in the topic, from researchers and students,
% to system architects and developers.

@@

\begin{bottomstuff}
Authors' addresses: Deprt. of Informatics and Telecommunications,
University of Athens, Athens, Greece. Deprt. of Software
Engineering, Izmir University of Economics, Izmir, Turkey.\\
emails: \{vsmoustakas,mema,ad\}@di.uoa.gr, huseyin.akcan@ieu.edu.tr
\end{bottomstuff}
